---
title: "Function to produce a single file from audio in"
author: "Spencer Graves"
date: 2018-10-28
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading a radio tuner 

I want to create *.wav files with names like "sourceYYYY_MM_DDTHH_MM.wav", where:  

* source = the name of an audio input, e.g., connected to the stereo mini phone jack for "audio in", which I can confirm the computer is digitizing by listening to something connected to "audio out" from that computer.  In my application, source = "KKFI" = the name of an FM radio station selected by a radio tuner.  If "audio in" port(s) do not work for this, try a USB Sound Adapter.  

* YYYY_MM_DDTHH_MM = the local [date and time of the start of that file following ISO 8601](https://en.wikipedia.org/wiki/ISO_8601).  At least initially, these will be 10 minute files staring on natural 10-minute breaks per the clock on the computer, which presumably is synced with [the official US time from the National Institute of Standards and Technology in Boulder, CO.](https://www.time.gov/) (I just confirmed that this time matches on both my Windows 7 desktop computer and my macOS 10.14 notebook.)

* ".wav" = the standard uncompressed "Waveform Audio File Format".  

At 10 minutes per file, this would create 6 files per hour and 144 such files in a 24-hour day.  The first such file created when the code starts will be short, because it will be set to end at exactly a 10-minute mark.  The software will be written to store a maximum of, say, N = 1440 such files, deleting the oldest when the max would otherwise be exceeded.  

I failed to find anything in R that could read an audio stream like this in real time.  

Fortunately, a post to Stack Overflow discussed [reading input sound signals using python](https://stackoverflow.com/questions/35344649/reading-input-sound-signal-using-python). This suggested the ["pyaudio"](https://people.csail.mit.edu/hubert/pyaudio/) package, which I installed from a Command Prompt as follows:  

> python -m pip install pyaudio

This Stack Overflow discussion included code in function "radioMonitor0_1" in "radioPackage\radioPy".

Let's first check the current working directory:  

```{python dir}
import os
cwd = os.getcwd()
print(cwd)
print(cwd.split('\\'))
```


```{r startTimer}
startPyTime <- proc.time()
```

The following returned "ModuleNotFoundError: No module named 'radioPackage':

#```{python pyaudio1}
import os
cwd = os.getcwd()
print(cwd)

from radioPackage.radioPy import radioMonitor0_1

radioMonitor0_1.radioMonitor0_1()
#```

However, the same script in file "idle0_1.py" was found and executed using a system command from within R: 

```{r sysPy0_1}
system("py idle0_1.py")
```

```{r etPy5}
(etPy5 <- proc.time()-startPyTime)
```

The "output.wav" file is 861 KB for 5 seconds of audio. 

```{r HDreqd}
(KBperMin <- 12*861)
```

So, a minute would require roughly 10 MB.  An hour would be roughly 600 MB, and a day would be:  

```{r GBreqd}
(KBperDay <- 24*60*KBperMin)/1e6
```

This means we need roughly 15 GB to store the audio from one day in this format.  Ten days would require roughly 150 GB.  I have roughly three times that much free space on the computer I plan to use for this.  I can do this. Later I may want to store only 5 days or 3 like this depending on how automated we might get with doing various kinds of spectral analysis on the fly and storing those summaries separate from the full detail in the "*.wav" format.  

